---
title: "Lab 5: Hypothesis Testing for Comparing Two Proportions / Means"
author: "Gabriel Etrata and Craig Balkhuysen"
output: html_document
date: April 14, 2016
---

<!-- WARNING: This file must be in the same folder with the data and source code to work. DO NOT MODIFY THE CODE CHUNK BELOW OR CHANGE ANY FILENAMES-->

```{r, echo=FALSE, message=FALSE}
#Source the source code
source("lab5-source-code.R")
#Import the data
lab5.data <- read.csv("lab5-data.csv", header = T)
#Attach the data
attach(lab5.data)
```
### Question 1
1-  
Is there a difference in how much Year-1 students exercise vs. Year-2 students?

2-
```{r}
compareMeans(hrs.exercise, year)
boxplot(hrs.exercise~year)
```

3-

$$H_o: \hat{p}_Y1 = \hat{p}_Y2$$

Null- There is no diffrence in the amount of hours exercised by Year 1s and Year 2s. 

$$H_a: \hat{p}_Y1 \neq \hat{p}_Y2$$

Alternative- There is a diffrence in the amount of hours exercised by Year 1s and Year 2s.

```{r}
lab5.data$shuffled.hrs.exercise <- sample(hrs.exercise)
compareMeans(lab5.data$shuffled.hrs.exercise, lab5.data$year)
```

$$\bar{x}_y1 - \bar{x}_y2 = -1.11$$

```{r}
set.seed(125)
randomization.test(lab5.data$shuffled.hrs.exercise, lab5.data$year, num = 1000)
```

Two Sided P-Value: 254


3-


4-

### Question 2

1-
 Is there a difference in how Year 1s and Year 2s feel about the Sophmore lunch issue?
 
2-
```{r}
compareProportions(soph.lunch, year)
table(soph.lunch, year)
```

 Null- There is no diffrence of opinion bewtween Year 1s and Year 2s on the sophmore lunch issue.
 
 $$H_a: \bar{x}_y1 = \bar{x}_y2$$
 
 Alternative- There is a difference of opinion between Year 1s and Year 2s on the sophmore lunch issue.
 
 $$H_o: \bar{x}_y1 \neq \bar{x}_y2$$
 
### Question 3

### Question 4
<!-- Continue your lab below. DON'T FORGET: before each use of the sample or randomization.test command, add set.seed() with some number of your choice in the parentheses. ALSO: when done, remove the "dist=T" argument from each randomization.test command in order to leave the long list of numbers out of your lab report.-->
